---
title: "Homework 10"
author: "Tobias Boggess"
date: '2022-05-01'
output: 
  pdf_document: 
    latex_engine: xelatex
---

```{r include = FALSE}
library(mdsr)
library(DBI)
library(RMySQL)
library(dplyr)
library(RSQLite)
library(dbplyr)
library(tidyr)
library(tidyverse)
```

## Chapter 15 Problems

### Problem 2:  Identify what years of data are available in the flights table of the airlines database.

Code:
```{r include = TRUE}
con <- dbConnect_scidb("airlines")
# con <- dbConnect(drv = MySQL(),
#                  dbname = "airlines",
#                  host = "mdsr.cdc7tgkkqd0n.us-east-1.rds.amazonaws.com",
#                  user = "mdsr_public",
#                  password = "ImhsmflMDSwR")

dbGetQuery(conn = con,
           statement = "SELECT year FROM flights
           GROUP BY year
           ORDER BY year")
```
The years available in the flights data are from 2010 to 2017. 
\newpage

### Problem 3: Use the dbConnect_scidb function to connect to the airlines database to answer the following problem. How many domestic flights flew into Dallas-Fort Worth (DFW) on May 14, 2010?

Code:
```{r include = TRUE}
dbGetQuery(conn = con,
           statement = "SELECT COUNT(*) AS count FROM flights
           WHERE dest = 'DFW' AND month = 5 AND day = 14 AND year = 2010;")
```
There are 754 flights that flew into Dallas Fort Worth on May 14, 2010. 

### Problem 5: Use the dbConnect_scidb function to connect to the airlines database to answer the following problem. Of all the destinations from Chicago O’Hare (ORD), which were the most common in 2010?

Code:
```{r include = TRUE}
dbGetQuery(conn = con,
           statement = "SELECT dest, origin, COUNT(*) AS count
           FROM flights
           WHERE year = 2010 AND origin = 'ORD'
           GROUP BY dest
           ORDER BY count DESC
           LIMIT 0,5;")
```
The most popular destination from Chicago O'Hare airport was LGA with 9787 flights.  
\newpage

### Problem 6: Use the dbConnect_scidb function to connect to the airlines database to answer the following problem. Which airport had the highest average arrival delay time in 2010?

Code:
```{r include = TRUE}
dbGetQuery(conn = con,
           statement = "SELECT dest, AVG(arr_delay) AS avg_arr_delay
           FROM flights
           WHERE year = 2010
           GROUP BY dest
           ORDER BY avg_arr_delay DESC
           LIMIT 0,5;")
```
GUM airport had the highest average arrival delay time in 2010. 

### Problem 8: Use the dbConnect_scidb function to connect to the airlines database to answer the following problem. List the airline and flight number for all flights between LAX and JFK on September 26th, 1990.

Code:
```{r include = TRUE}
# dbGetQuery(conn = con,
#            statement = "(SELECT carrier, flight 
#            FROM flights
#            WHERE origin = 'LAX' AND dest = 'JFK'
#            AND year = 1990 AND month = 9 AND day = 26)
#            UNION
#            (SELECT carrier, flight
#            FROM flights
#            WHERE origin = 'JFK' AND dest = 'LAX'
#            AND year = 1990 AND month = 9 AND day = 26);")

dbGetQuery(conn = con,
           statement = "SELECT carrier, flight 
           FROM flights
           WHERE origin IN ('LAX', 'JFK') AND 
           dest IN ('LAX', 'JFK') AND year = 1990 AND
           month = 9 AND day = 26;")
```
The above doesn't return any flights because the flights dataset is only available between 2010 and 2017. 

```{r include = FALSE}
dbDisconnect(conn = con)
```









