---
title: "Class_Exercises_ClassNotes_3"
author: "Tobias Boggess"
date: "2/14/2022"
output: pdf_document
---
## Section 4.2 Exercises

### Exercise 1: Show the following

#### a) select(.data = flights, year, day)
\
Results:
```{r include = TRUE}
library(nycflights13)
library(dplyr)
library(ggplot2)
select(.data = flights, year, day)
```
\
My guess: Will return the data from columns year and day.
\newpage

#### b) select(.data = flights, year:day)
\
Results:
```{r include = TRUE}
select(.data = flights, year:day)
```
\
My Guess: This will return all columns from year to day.

#### c) select(.data = flights, -(year:day))
\
Results
```{r include = TRUE}
select(.data = flights, -(year:day))
```
\
My Guess: All columns will be shown except the ones from year to day.
\newpage

### Exercise 2: Guess and check.
\
Code:
```{r include = TRUE}
names(flights)
```

#### a) select(.data = flights, starts_with("sched"))
\
Results:
```{r include = TRUE}
select(.data = flights, starts_with("sched"))
```
\
My Guess: This will show columns that have a header matching "sched".
\newpage

#### b) select(.data = flights, contains("arr"))
\
Results:
```{r include = TRUE}
select(.data = flights, contains("arr"))
```
\
My Guess: This will show columns containing the string "arr". 

#### c) select(.data = flights, starts_with("dep_"), starts_with("arr_"))
\
Results:
```{r include = TRUE}
select(.data = flights, starts_with("dep_"), starts_with("arr_"))
```
\
My Guess: This will show all the columns that start with "dep_" or "arr_". 
\newpage

## Section 4.4 Exercises

### Exercise 3: Report R commands that use filter() and the logical operators (’&’, ’|’, and ’!’) with the flights data to find all flights that:

* a) Had an arrival delay of two or more hours.
* b) Flew to Houston (IAH or HOU).
* c) Were operated by United, American, or Delta.
* d) Departed in summer (July, August, or September).
* e) Departed between midnight and 6:00 AM (inclusive).
* f) Were operated by United, departed in July, and had an arrival delay of two or more hours.
\
Code:
```{r include = TRUE}
filter(.data = flights, arr_delay >= 120)
filter(.data = flights, dest == "IAH" | dest == "HOU")
filter(.data = flights,
       carrier == "UA" |
         carrier == "AA" | carrier == "DL")
filter(.data = flights, month == 7 | month == 8 | month == 9)
```
\newpage
```{r include = TRUE}
filter(.data = flights, dep_time == 2400 | dep_time <= 600)
filter(.data = flights, carrier == "UA" &
         month == 7 & arr_delay >= 120)
```
\newpage

## Section 4.5 Exercises

### Exercise 4: Report R commands that use arrange() to sort the flights data to:

* a) Find the flights that had the shortest delays.
* b) Find the flights that had the longest delays.
* c) Find the flights that had the earliest departure times.
* d) Find the flights that had the latest departure times.
* e) Find the flights that traveled the shortest distance.
* f) Find the flights that traveled the longest distance.
\
Code:
```{r include = TRUE}
arrange(.data = flights, dep_delay)
arrange(.data = flights, desc(dep_delay))
arrange(.data = flights, dep_time)
arrange(.data = flights, desc(dep_time))
```
\newpage
```{r include = TRUE}
arrange(.data = flights, distance)
arrange(.data = flights, desc(distance))
```
\newpage

### Exercise 5: Do the following.
\
Given Data Frame:
```{r include = TRUE}
x <- data.frame(
  x1 = c(2, 1, NA, 8, 7, 5, 4),
  x2 = c("a", NA, "c", "d", "c", "a", "d"),
  stringsAsFactors = FALSE
)
x
```

#### a) Guess what the following will do.
\
Code: 
```{r include = TRUE}
arrange(.data = x, is.na(x1))
```
\
My Guess: This will arrange the results so na is at the bottom of the data frame.
\newpage

#### b) Guess what the following will do.
\
Code:
```{r include = TRUE}
arrange(.data = x, desc(is.na(x1)))
```
\
My Guess: This will arrange all the values that are na at the top based on the x1 column. 

## Section 4.6 Exercises

### Exercise 6: Report an R command that uses mutate() or transmute(), with flights, to compute arr_time - dep_time, and compare it with air_time. Why are they different?
\
Code: 
```{r include = TRUE}
mutate(.data = flights, new_air_time = arr_time - dep_time)
all(flights$air_time == flights$new_air_time)
```
\
They are different and its because one uses hours and minutes and the other air_time variable is in minutes.

### Exercise 7: Execute the code.
\
Code:
```{r include = TRUE}
flights_small <- select(.data = flights,
                        year:day,
                        ends_with("delay"),
                        distance,
                        air_time)
mutate(
  .data = flights_small,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)
```
\
The gain_per_hour does seem to get computed.
\newpage

## Section 4.7 Exercises

### Exercise 8: Use rename() to change the names of the variables in z to new_z1, new_z2, and new_z3. Report your R command(s).
\
Code: 
```{r include = TRUE}
z <- data.frame(
  z1 = c(5, 4, 3),
  z2 = c("a", "c", "b"),
  z3 = c(14, 22, 13)
)
new_z <- rename(
  .data = z,
  new_z1 = z1,
  new_z2 = z2,
  new_z3 = z3
)
z
new_z
```
\newpage

## Section 4.8 Exercises

### Exercise 9: Create the not_cancelled data frame (using flights):
\
Code:
```{r include = TRUE}
not_cancelled <-
  filter(.data=flights,
         !is.na(dep_delay),
         !is.na(arr_delay)
         )
```

#### a) Use summarize() with median() to find the median departure delay and the median arrival delay. Report the two values.
\
Code:
```{r include = TRUE}
summarize(.data = not_cancelled,
          med_dep_delay = median(dep_delay),
          med_arr_delay = median(arr_delay))
```

#### b) Use summarize() with max() to find the longest departure delay and the longest arrival delay.Report the two values.
\
Code: 
```{r include = TRUE}
summarize(.data = not_cancelled,
          long_dep_delay = max(dep_delay),
          long_arr_delay = max(arr_delay))
```
\newpage

#### c) Use summarize() with min() to find the shortest departure delay and the shortest arrival delay.Report the two values.
\
Code:
```{r include = TRUE}
summarize(.data = not_cancelled,
          short_dep_delay = min(dep_delay),
          short_arr_delay = min(arr_delay))
```

### Exercise 10: Answer the following.

#### a) What does the following command do?
\
Code:
```{r include = TRUE}
summarize(.data = not_cancelled,
          total_flights = n())
```
\
This shows the total number of flights in the flights data set.

#### b) What does the following command do?
\
Code:
```{r include = TRUE}
summarize(.data = not_cancelled,
          hour_arr_delay_total = sum(arr_delay > 60))
```
\
This will show the number of times the arrival delay was greater than an hour. 
\newpage

#### c) What does the following command do?
\
Code:
```{r include = TRUE}
summarize(.data = not_cancelled,
          hour_arr_delay_proportion = 
            sum(arr_delay > 60) / n()
          )
```
\
This will show the proportion of times a flight was delayed by more than an hour by the number of flights.

## Section 4.9 Exercises

### Exercise 11: Do the following.

#### a) Explain in words what the following commands do (recall that dest is the destination of the flight):
\
Code: 
```{r include = TRUE}
by_dest <- group_by(.data = flights, dest)
delay_by_dest <- summarize(.data = by_dest,
                           mean_arr_delay = mean(arr_delay, na.rm = TRUE))
delay_by_dest
```
\
The above code calculates the mean arrival delay of every destination airport flown to from New York. 
\newpage

#### b) summarize() can summarize more than one variable at a time. Explain in words what the following commands do:
\
Code:
```{r include = TRUE}
by_dest <- group_by(.data = flights, dest)
delay_dist_by_dest <- summarize(
  .data = by_dest,
  mean_dist = mean(distance, na.rm = TRUE),
  mean_arr_delay = mean(arr_delay, na.rm = TRUE)
)
delay_dist_by_dest
```
\
Summarize can do more than one variable at a time. The output of them is the mean of the distance traveled between a New York airport and the destination airport. The second variable is the mean distribution of the arrival delay at each airport. 
\newpage

### Exercise 12: Do the following.
\
Code:
```{r include = TRUE}
resp <- c(23, 11, 14, 16, 19, 26, 24, 29, 31, 28, 34, 25)
trt <- c(rep("Ctrl", 4), rep("TrtA", 4), rep("TrtB", 4))
age <- c(33, 45, 30, 24, 22, 31, 39, 40, 29, 19, 27, 25)
gndr <- c("m", "m", "f", "f", "m", "f", "f", "m", "f", "m", "f", "m")
ExpData <- data.frame(
  TrtGrp = trt,
  SubjectGender = gndr,
  SubjectAge = age,
  Response = resp,
  stringsAsFactors = FALSE
)
```

#### a) Explain in words what the following commands do:
\
Code: 
```{r include = TRUE}
by_TrtGrp <- group_by(.data = ExpData, TrtGrp)
summarize(.data = by_TrtGrp, Count = n())
```
\
The command will show how many of each group there are.

#### b) Use group_by() and summarize() to compute the mean Response by TrtGrp. Report the three
mean Response values.
\
Code:
```{r include = TRUE}
TrtGroupBy <- group_by(.data = ExpData, TrtGrp)
summarize(.data = TrtGroupBy, meanResponse = mean(Response, na.rm = TRUE))
```

#### c) Now use group_by() and summarize() to compute the mean Response and mean SubjectAge by TrtGrp. Report the three mean Response values and the three mean SubjectAge values.
\
Code:
```{r include = TRUE}
TrtGroupBy <- group_by(.data = ExpData, TrtGrp)
summarize(
  .data = TrtGroupBy,
  meanResponse =
    mean(Response, na.rm = TRUE),
  meanSubjectAge = mean(SubjectAge, na.rm = TRUE)
)
```

### Exercise 13: Do the following with the flights data set.

#### a) Use group_by(), summarize() with n(), and arrange() to determine which tailnum (i.e. which individual airplane) flew the most times. Report the tailnum value of the airplane.
\
Code:
```{r include = TRUE}
tailNumGroup <- group_by(.data = flights, tailnum)
mostTailNum <- summarize(.data = tailNumGroup, Count = n())
arrange(.data = mostTailNum, desc(Count))
```
\newpage

#### b) Use group_by(), summarize() with n(), and arrange() to determine which dest (i.e. which destination) was flown to the most times. Report the (abbreviated) destination name.
\
Code:
```{r include = TRUE}
destGroup <- group_by(.data = flights, dest)
mostDestGroup <- summarize(.data = destGroup, Count = n())
arrange(.data = mostDestGroup, desc(Count))
```

## Section 4.10 Exercises

### Exercise 14: Do the following with the command below.
\
Code:
```{r include = TRUE}
x <- c(2, 5, 4, 3, 7, 9)
```

#### a) What does the following command do?
\
Code:
```{r include = TRUE}
x %>% mean()
mean(x)
```
\
This is a way to apply x to the mean function.  
\newpage

#### b) What does the following command do?
\
Code:
```{r include = TRUE}
x %>% mean() %>% sqrt() %>% round(digits = 2)
round(sqrt(mean(x)), digits = 2)
```
\
This will pass x into the mean() function and from there the result will be passed to the square root function and the result of that will be passed to the round function. 

#### c) Rewrite the following sequence of commands using the pipe operator.
\
Original Code:
```{r include = TRUE}
mean_x <- mean(x)
sqrt_mean_x <- sqrt(mean_x)
round_sqrt_mean_x <- round(sqrt_mean_x, digits = 2)
round_sqrt_mean_x
```
\
Modified Code:
```{r include = TRUE}
x %>% mean() %>% sqrt() %>% round(digits = 2)
```

#### d) Rewrite the following command using the pipe operator. Report your R command(s).
\
Original Code:
```{r include = TRUE}
round(sqrt(mean(x)), digits = 2)
```
\
Modified Code:
```{r include = TRUE}
x %>% mean() %>% sqrt() %>% round(digits = 2)
```
\newpage

### Exercise 15: Do the following.
#### a) Rewrite the following command using the pipe operator.
\
Command:
```{r include = TRUE}
delay <- select(.data = flights, arr_delay)
```
\
Pipe Operator version:
```{r include = TRUE}
flights %>% select(arr_delay)
```

#### b) Rewrite the following command using the pipe operator.
\
Original:
```{r include = TRUE}
dest_delay <- select(.data = flights, dest, arr_delay)
```
\
Pipe Operator version:
```{r include = TRUE}
flights %>% select(dest, arr_delay)
```

#### c) Rewrite the following pair of commands using the pipe operator.
\
Original:
```{r include = TRUE}
dest_delay <- select(.data = flights, dest, arr_delay)
sea_den <- filter(.data = dest_delay,
                  dest == "SEA" | dest == "DEN")
sea_den
```
\
Pipe Operator version:
```{r include = TRUE}
flights %>% select(dest, arr_delay) %>% filter(dest == "SEA" | dest == "DEN")
```
\newpage

#### d) Rewrite the following sequence of commands using the pipe operator.
\
Original:
```{r include = TRUE}
dest_delay <- select(.data = flights, dest, arr_delay)
sea_den <- filter(.data = dest_delay,
                  dest == "SEA" | dest == "DEN")
by_dest <- group_by(sea_den, dest)
delay_by_dest <- summarize(by_dest,
                           mean_arr_delay = mean(arr_delay, na.rm = TRUE))
delay_by_dest
```
\
Pipe Operator Version:
```{r include = TRUE}
flights %>% select(dest, arr_delay) %>% filter(dest == "SEA" |
                                                 dest == "DEN") %>% group_by(dest) %>%     summarize(mean_arr_delay = mean(arr_delay, na.rm = TRUE))
```
\newpage

### Exercise 16: Rewrite the following command using the pipe operator.
\
Original Code:
```{r include = TRUE}
den_delays <- summarize(
  filter(.data = flights,
         dest == "DEN"),
  mean_dep_delay = mean(dep_delay, na.rm = TRUE),
  mean_arr_delay = mean(arr_delay, na.rm = TRUE)
)
den_delays
```
\
Revised Code:
```{r include = TRUE}
flights %>% filter(dest == "DEN") %>% summarize(
  mean_dep_delay = mean(dep_delay, na.rm = TRUE),
  mean_arr_delay = mean(arr_delay, na.rm = TRUE)
)
```
\newpage

## Section 4.11 Exercises
### Exercise 17: Do the following with the given data frames.
\
Data frames:
```{r include = TRUE}
df1 <- data.frame(Respondent_ID = c(1001, 1002, 1003),
                  Q1_Response = c(55, 62, 39))
df1

df2 <- data.frame(
  Respondent_ID = c(1002, 1003, 1004),
  Q2_Response = c("yes", "no", "yes")
)
df2
```

#### a) Guess what the result of the following command will be, then check your answer and report the result.
\
Code:
```{r include = TRUE}
inner_join(x = df1, y = df2, by = "Respondent_ID")
```
\
This code will display the matching rows in both data frames based on the Respondent ID. 
\newpage

#### b) Guess what the result of the following command will be, then check your answer and report the result.
\
Code:
```{r include = TRUE}
left_join(x = df1, y = df2, by = "Respondent_ID")
```
\
This will show all the rows from the df1 data frame and include information about the Q2_response for al variables in df1. 

#### c) Guess what the result of the following command will be, then check your answer and report the result.
\
Code:
```{r include = TRUE}
full_join(x = df1, y = df2, by = "Respondent_ID")
```
\
This code will show all rows for both df1 and df2 even if df1 doesn't have anything for the column containing df2. 

#### d) If we didn’t specify by = "Respondent_ID", by default what key variable would each of the *_join() functions use to match rows?
\
Code:
```{r include = TRUE}
full_join(x = df1, y = df2)
```
\
The default is the first column of df1. 
\newpage

#### e) What would happen if Q1_Response and Q2_Response were both named Response in the two data frames, e.g.
\
Code:
```{r include = TRUE}
#df1 <- rename(.data = df1, Response = Q1_Response)
#df2 <- rename(.data = df2, Response = Q2_Response)
#full_join(x = df1, y = df2)
```
\
This will result in an error "Joining, by = c("Respondent_ID", "Response")
Error in `full_join()`:
! Can't join on `x$Response` x `y$Response` because of incompatible
  types.
i `x$Response` is of type <double>>.
i `y$Response` is of type <character>>.
Backtrace:
 1. dplyr::full_join(x = df1, y = df2)
 2. dplyr:::full_join.data.frame(x = df1, y = df2)
 Error in full_join(x = df1, y = df2) : 
i `x$Response` is of type <double>>.
i `y$Response` is of type <character>>.".
\

#### f) What would happen if, as in part e, Q1_Response and Q2_Response were both named Response, and we typed:
\
Code:
```{r include = TRUE}
#df1 <- rename(.data = df1, Response = Q1_Response)
#df2 <- rename(.data = df2, Response = Q2_Response)
#inner_join(x = df1, y = df2)
```
\
This reports an error, "Error in `stop_subscript()`:
! Can't rename columns that don't exist.
x Column `Q1_Response` doesn't exist.
Backtrace:
  1. dplyr::rename(.data = df1, Response = Q1_Response)
  2. dplyr:::rename.data.frame(.data = df1, Response = Q1_Response)
  3. tidyselect::eval_rename(expr(c(...)), .data)
  4. tidyselect:::rename_impl(...)
  5. tidyselect:::eval_select_impl(...)
     ...
 20. tidyselect:::chr_as_locations(x, vars)
 21. vctrs::vec_as_location(x, n = length(vars), names = vars)
 22. vctrs `<fn>`()
 23. vctrs:::stop_subscript_oob(...)
 24. vctrs:::stop_subscript(...)
 Error in stop_subscript(class = "vctrs_error_subscript_oob", i = i, subscript_type = subscript_type, :".
 
\newpage

### Exercise 18: Do the following with the given data frames. 
\
Data Frames:
```{r include = TRUE}
df1 <- data.frame(Respondent_ID = c(1000, 1001, 1002, 1003, 1004, 1005, 1006),
Q1_Response = c(55, 62, 39, 45, 70, 77, 56))
df1

df2 <- data.frame(Respondent_ID = c(1003, 1002, 1000, 1004, 1006, 1001, 1005),
Q2_Response = c(12, 17, 23, 24, 19, 30, 20))
df2
```

#### a) What happens to the ordering of the rows of df2 when you combine it with df1 using:
\
Code: 
```{r include = TRUE}
inner_join(x = df1, y = df2, by = "Respondent_ID")
```
\
The ordering will be based on the Respondent ID in df1. 
\newpage

#### b) How would the result differ if you swapped the roles of df1 and df2, e.g.:
\
Code:
```{r include = TRUE}
inner_join(x = df2, y = df1, by = "Respondent_ID")
```
\
The code changes so the Respondent ID will be based on df2 instead of df1. 

### Exercise 19: Do the following with the given data frames. 
\
Data frames:
```{r include = TRUE}
dfX <- data.frame(
  LastName = c("Smith", "Smith", "Jones", "Smith",
               "Olsen", "Taylor", "Olsen"),
  FirstName = c("John", "Kim", "John", "Marge", "Bill",
                "Bill", "Erin"),
  Gender = c("M", "F", "M", "F", "M", "M", "F"),
  ExamScore = c(75, 80, 64, 78, 90, 89, 79)
)
dfX
```
\newpage
```{r include = TRUE}
dfY <- data.frame(
  LastName = c("Olsen", "Jones", "Taylor", "Smith",
               "Olsen", "Smith", "Smith"),
  FirstName = c("Bill", "John", "Bill", "Kim", "Erin",
                "John", "Marge"),
  Gender = c("M", "M", "M", "F", "F", "M", "F"),
  Grade = c("A", "D", "B", "B", "C", "C", "C")
)
dfY
```

#### a) Write a command involving, say, full_join() that combines the two data frames by person. 
\
Code:
```{r include = TRUE}
full_join(x = dfX, y = dfY, by = c("LastName", "FirstName"))
```
\newpage

#### b) What happens with the third variable (Gender) when you only specify the other two (LastName and FirstName) as the key via the by argument? 
\
Code:
```{r include = TRUE}
full_join(x = dfX, y = dfY, by = c("LastName", "FirstName", "Gender"))
full_join(x = dfX, y = dfY)
full_join(x = dfX, y = dfY, by = c("LastName", "FirstName"))
```
\newpage

#### c) What would happen if you tried to combine dfX and dfY only specifying LastName as the key variable?
\
Code:
```{r include = TRUE}
full_join(x = dfX, y = dfY, by = "LastName")
```