---
title: "Classnotes 8"
author: "Tobias Boggess"
date: "4/4/2022"
output: 
  pdf_document: 
    latex_engine: xelatex
---
```{r include = FALSE, echo = FALSE}
library(ggplot2)
```


## Section 13.2 Exercises

### Exercise 1: Do the following.

#### a) Use set.seed() to set the ”seed” to a value (any positive integer will do). Then use sample() to generate n = 5 random numbers from 1, 2, . . . , 100. Report your R commands.
\
Code:
```{r include = TRUE}
set.seed(5)

sample(x = 1:100, size = 5)
```

#### b) Now set the ”seed” again (to the same value), then use sample() again to produce n = 5 random numbers from 1, 2, . . . , 100. Confirm that this regenerates the same five values you got in part a.
\
Code:
```{r include = TRUE}
set.seed(5)

sample(x = 1:100, size = 5)
```

#### c) What would’ve happened in part b if you hadn’t set the ”seed” prior to the call to sample()? Try it.
\
Code:
```{r include = TRUE}
set.seed(23)

sample(x = 1:100, size = 5)
```
I would have received different numbers since the seed is different. 

### Exercise 2: Do the following with uniform random variables.

#### a) Use runif() to generate n = 1,000 random values between 0 and 1. Save them in a vector x. Report your R command(s).
\
Code:
```{r include = TRUE}
set.seed(500)

x <- runif(n = 1000, min = 0, max = 1)
head(x, n = 5)
```

#### b) Produce a histogram of the simulated data:Are the simulated values fairly evenly spread over the interval from 0 to 1?
\
Code:
```{r include = TRUE}
ggplot(data.frame(x = x), mapping = aes(x = x)) +
  geom_histogram(binwidth = 0.1,
                 boundary = 0.0,
                 fill = "blue")
```

Overall, the values are fairly evenly spread across the interval between 0 and 1 according to the histogram. 

### Exercise 3: Do the following with normal random numbers.

#### a) Use rnorm() to generate n = 1,000 random values from the normal(μ,σ) curve, with with μ = 0 and σ = 1. Save them in a vector x. Report your R command(s).
\
Code:
```{r include = TRUE}
set.seed(21)

x <- rnorm(n = 1000, mean = 0, sd = 1)
head(x, n = 5)
```

#### b) Produce a histogram of the simulated data. Do the simulated values follow a (approximately) bell-shaped pattern from -3 to 3?
\
Code:
```{r include = TRUE}
ggplot(data.frame(x = x), mapping = aes(x = x)) +
  geom_histogram(binwidth = 0.5,
                 boundary = -3.5,
                 fill = "blue")
```
The histogram does show a bell-shaped pattern from -3 to 3. 

### Exercise 4: Do the following with dichotomous random numbers.

#### a) Use rbinom() to generate a dichotomous (0 or 1) sequence of ten (n = 10) ”flips” of a biased ”coin” that lands ”heads” (i.e. results in a 1) with probability 0.7 (prob = 0.7) and ”tails” (0) otherwise. Report your R command(s).
\
Code:
```{r include = TRUE}
set.seed(64)

rbinom(n = 10, size = 1, prob = 0.7)
```

#### b) Now use rbinom() to generate a sequence of ten (n = 10) dichotomous outcomes that are decreasingly likely be 1 according to the following probabilities. Report your R command(s).
\
Code:
```{r include = TRUE}
# set.seed(56)
my.probs <- seq(from = 0.95, to = 0.05, by = -0.1)

rbinom(n = 10, size = 1, prob = my.probs)
```

## Section 13.3 Exercises

### Exercise 5: This exercise involves simulating data from a logistic regression model to investigate the performance of parameter estimates, the (estimated) intercept b0 and slope b1.

#### a) generate n = 1,000 dichotomous observations from a logistic regression model, with (true) parameter values intercept β0 = 4 and slope β1 = −1, at values of the explanatory variable generated from a uniform(0, 10) distribution. Report your R commands.
\
Code:
```{r include = TRUE}
set.seed(57)

r <- runif(n = 1000, min = 0, max = 10)

true_probs <- exp(4 + (-1 * r)) / (1 + exp(4 + (-1 * r)))

y <- rbinom(n = 1000, size = 1, prob = true_probs)

sim.data <- data.frame(X = r, Y = y)
head(sim.data)
```

#### b) Fit a logistic regression model to your simulated data, and report the resulting parameter estimates b0 and b1 from the output of summary(). Are they close to the true parameter values β0 = 4 and β1 = −1?
\
Code:
```{r include = TRUE}
logreg <- glm(Y ~ X, data = sim.data, family = "binomial")
summary(logreg)
```

The values from the summary of the logistic model are 4.17255 for $\beta_0$ and -1.03260 for $\beta_1$.

